---
title: "Covid91 vaccine study FinalFall2022"
author: "John Garza"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="finalFall2022.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
```

# Introduction

A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

## Males

```{r}
Males_data <- filter(FinalData, sex == "M")
```

```{r}
barchartGC(~infected + treatment,data=Males_data)
barchartGC(~infected + treatment,data=Males_data , type="percent", main = ("Infected vs Treatment"))
```

Male cases from the clinical trials are in the barplot's above. The blue bar's are the males who were given a plecebo or fake medications and the pink bar's are the males who were given a vaccine or an actual strong medication.    

### Numerical results

```{r}
table1 <- xtabs(~infected + treatment, data=Males_data)
rowPerc(table1)
colPerc(table1)
```

The male's numerical results displays the statistical results for an infected, ok, and vaccinated or not vaccinated male. On the fist table it reads that there are 68 males that are sick but are given a plecebo or fake medications out of 100 males and 32 males who were given a vaccine out of the 100 males. Then the table below has the results for the trial. The chance for a positive covid19 test under (placebo) = 2.7 , and the chance for a positive covid19 test under (vaccine) = 1.2, OK patients are negative tests and are (placebo) = 97.3 and (vaccine) = 98.8. Men who took the vaccine are safer than men who did not in this particular set.

### Inferential 


```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
malesfish <- fisher.test(table1)
malesfish
```

The information we are generally looking for here is the Fishers Exact test. In this scenario the test results read p-value = 2e-12, alternative hypothesis: true odds ratio is not equal to 1, 95 percent confidence interval: 1.7 2.8, sample estimates: odds ratio 2.2. Going from these results we can assume that there is no association between the effectiveness across the sub population tracked because the p value is smaller than the significant 0.05. 


## Effectiveness Confidence Interval 'r low' to 'r hi'


```{r}
effect <- malesfish$p.value/(malesfish$p.value +1)
effectivenessmales = malesfish$estimate/(malesfish$estimate +1)*100
print(effectivenessmales)
```

```{r}
low=(malesfish$conf.int[1]/(malesfish$conf.int[1] +1))*100
hi=(malesfish$conf.int[2]/(malesfish$conf.int[2] +1))*100
options(digits=2)
```

This odds ratio indicates the effectiveness of the shot given to a group of males from 18,125 observations. The chance of this shot being effective for males is 69, but there is not a supported background to this result. For example the previous barplot, numerical results, and the inferential Fisher's Exact Test for Count Data.
    

## Females

```{r}
females_data <- filter(FinalData, sex == "F")
```

```{r}
barchartGC(~infected + treatment,data=females_data)
barchartGC(~infected + treatment,data=females_data , type="percent", main = ("Infected vs Treatment"))
```

Female cases from the clinical trials are in the barplot's above. The outcome for females is different than the males barchart, but there is still a group of females who are sick and one group who are ok. Then the blue bar's are the females who were given a placebo or fake medications and the pink bar's are the females who were given a vaccine or an actual strong medication and still got covid on the left or was somehow prevented on the right.  

### Numerical results

```{r}
table1 <- xtabs(~infected + treatment, data=females_data)
rowPerc(table1)
colPerc(table1)
```

The female's numerical results displays the statistical results for the infected(vaccinated), infected(unvaccinated), OK(vaccinated), and OK(unvaccinated). On the fist table it reads that there are 43 females who went through a placebo effect and 57 females who were given a vaccine out of 100 females that got sick and 50 placebo and 50 vaccine females who are ok. Then the table below has the results for the trial. The chance for a positive covid19 test on covid(placebo) = 1.7 , and the chance for a positive covid19 test on covid(vaccine) = 2.3, OK patients are negative tests and resulted as OK = 98.3 under placebo and OK = 97.7 under vaccine. There was some how more females who were some how safe after under going a placebo effect so the vaccine was considered less valuable than the placebo in this particular set.   

### Inferential 


```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
femalesfish <- fisher.test(table1)
femalesfish
```

Interestingly similar to the males results the information we are generally looking for here is the Fishers Exact test. In this scenario the test results read p-value = 0.005, alternative hypothesis: true odds ratio is not equal to 1, 95 percent confidence interval: 0.60 0.92, sample estimates: odds ratio 0.74, and going from these results we can assume that there is also no association between the effectiveness across the sub population tracked.  

## Effectiveness Confidence Interval 'r low' to 'r hi'


```{r}
effect <- femalesfish$p.value/(femalesfish$p.value +1)
effectivenessfemales = femalesfish$estimate/(femalesfish$estimate +1)*100
print(effectivenessfemales)
```

```{r}
low=(femalesfish$conf.int[1]/(femalesfish$conf.int[1] +1))*100
hi=(femalesfish$conf.int[2]/(femalesfish$conf.int[2] +1))*100
options(digits=2)
```

In general, an odds ratio indicates a strong association. Since this number is smaller than the males odds then there is a possibility that the vaccine is not effective for females. That is why it appears to be that there is a significant association between effectiveness across the covid-19 virus. For example, in the females analysis results have gone to prove that we don't have an effective association in preventing covid for females.  

## LGBTQ

```{r}
LGBTQ_data <- filter(FinalData, LGBTQ == "gay")
```

```{r}
barchartGC(~infected + treatment,data=LGBTQ_data)
barchartGC(~infected + treatment,data=LGBTQ_data , type="percent", main = ("Infected vs Treatment"))
```

The barplot's above here shows the percent of LGBTQ individuals who were infected with covid19 and who were OK. The pink bars are people who received the vaccine and the blue bars are the people who received the placebo. In this barchart there was a significantly higher reported number of covid19 patients who received a vaccine over the covid19 patients who received a placebo and then there was a slightly bigger number of reported patients who are OK but received a vaccine over ones who were OK and received the placebo.  

### Numerical results

```{r}
table1 <- xtabs(~infected + treatment, data=LGBTQ_data)
rowPerc(table1)
colPerc(table1)
```

The LGBTQ numerical results displays the statistical results for an infected, ok, and vaccinated or not vaccinated LGBTQ person. On the fist table it reads that there are 13 people who went through a placebo effect and 87 people who were given a vaccine out of 100 people that are sick and 49 placebo and 51 vaccine people who are OK. Then the table below has the results for the bottom table that are covid19(placebo) = 0.54, covid19(vaccine) = 3.3, OK(placebo) = 99.46, and OK(vaccine) = 96.7. In this particular set they were not getting protected, there is a 3.3 chance of getting covid even after getting the vaccine.  

### Inferential 


```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
LGBTQfish <- fisher.test(table1)
LGBTQfish
```    

Interestingly in this scenario the test results read p-value = 1e-08, alternative hypothesis: true odds ratio is not equal to 1, 95 percent confidence interval: 0.066 0.344, sample estimates: odds ratio 0.16, going from these results we can also assume that there is also no association between the effectiveness across the variety of sub populations tracked. The chance that the vaccine was effective in protecting LGBTQ people from covid is slightly less than the chance it was for the vaccine's effectiveness in protecting males from covid, with the lgbtq value at 13.9 and the males value at 17.2. The vaccine is not likely to be effective in both future clinical trials.  

## Effectiveness Confidence Interval 'r low' to 'r hi'


```{r}
effect <- LGBTQfish$p.value/(LGBTQfish$p.value +1)
effectivenessLGBTQ = LGBTQfish$estimate/(LGBTQfish$estimate +1)*100
print(effectivenessLGBTQ)
```

```{r}
low=(LGBTQfish$conf.int[1]/(LGBTQfish$conf.int[1] +1))*100
hi=(LGBTQfish$conf.int[2]/(LGBTQfish$conf.int[2] +1))*100
options(digits=2)
```

In general, an odds ratio indicates a strong association. Since this ratio is 14 it is a small possibility, or not likely. That is why it appears to be that there is no significant association between the vaccine and the LGBTQ analysis.  

## Druggies

```{r}
druggies_data <- filter(FinalData, DrugUser == "yes")
```

```{r}
barchartGC(~infected + treatment,data=druggies_data)
barchartGC(~infected + treatment,data=druggies_data , type="percent", main = ("Infected vs Treatment"))
```

The barplot's above here shows the percent of drug user individuals who were infected with covid19 and who were OK. The pink bars are people who received the vaccine and the blue bars are the people who received the placebo. In this barchart there was a much greater reported number of covid19 patients who received a vaccine over the covid19 patients who received a placebo and then there was a slightly bigger number of reported patients who are OK but received a vaccine over ones who were OK and received the placebo.  

### Numerical results

```{r}
table1 <- xtabs(~infected + treatment, data=druggies_data)
rowPerc(table1)
colPerc(table1)
```

The drug using numerical results displays the statistical results for an infected, ok, and vaccinated or not vaccinated drug using person. On the fist table it reads that there are 17 people who went through a placebo effect and 83 people who were given a vaccine out of 100 people that are sick and 49 placebo and 51 vaccine people who are OK, like the LGBTQ results. Then the table below has the results for the bottom table that are covid19(placebo) = 0.81, covid19(vaccine) = 3.8, OK(placebo) = 99.19, and OK(vaccine) = 96.2. In this particular set of the trial the vaccine was not effective for covid patients and more effective for patients who tested negative.  

### Inferential 


```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
druggiesfish <- fisher.test(table1)
druggiesfish
```   

Interestingly in this scenario the test results read p-value = 3e-07, alternative hypothesis: true odds ratio is not equal to 1, 95 percent confidence interval: 0.094 0.418, and the sample estimates: odds ratio is 0.21. The values above provide a representation of the count for the drug using individuals in the clinical trial data.   

## Effectiveness Confidence Interval 'r low' to 'r hi'


```{r}
effect <- druggiesfish$p.value/(druggiesfish$p.value +1)
effectivenessdruggies = druggiesfish$estimate/(druggiesfish$estimate +1)*100
print(effectivenessdruggies)
```

```{r}
low=(druggiesfish$conf.int[1]/(druggiesfish$conf.int[1] +1))*100
hi=(druggiesfish$conf.int[2]/(druggiesfish$conf.int[2] +1))*100
options(digits=2)
```

In general, an odds ratio indicates a strong association. Since this ratio is 17 it is smaller than the previous odds ratio's, the vaccine had a small effective rate in preventing covid for drug user's.  


# Overall Results and Conclusions

After investigating the trials of a new  preventative vaccine. I learned that the null hypothesis was that the vaccine was effective in preventing infection from the covid19 epidemic across a variety of sub populations tracked and the alternative hypothesis was that the vaccine has exhibited different effectiveness across the variety of sub populations tracked. The first group was the males. After investigating how to build models that produced the outcomes needed I made an assessment based off of these outcomes. Based off of the Fisher's Exact Test for the male Count Data the vaccine is not effective and reject the null hypothesis and accept the alt hypothesis that is that the vaccine has exhibited different effectiveness across this variety of the sub populations tracked. Interestingly for every sub population tracked further investigation resulted in the outcome of rejecting the null hypothesis and accepting the alternative hypothesis. Based off of the models built I learned this, there is strong evidence supporting the alternative hypothesis in these models. For example the automated response for the P value in the sub groups tracked is, male p-value is 2e-12, female p-value is 0.005, LGBTQ p-value is 1e-08, and the drug user p-value is 3e-07. A statistically significant test result (P ≤ 0.05) means that the test hypothesis is false or should be rejected. After going through every sub population the outcome for the p value was a number extremely small and less than the 0.05 significant result. Therefor the vaccine should not be recommended, as it stands in these clinical trails.