---
title: "Covid91 vaccine study FinalFall2022"
author: "John Garza"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="finalFall2022.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
```

# Introduction

A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

## Males

```{r}
Males_data <- filter(FinalData, sex == "M")
```

```{r}
barchartGC(~infected + treatment,data=Males_data)
barchartGC(~infected + treatment,data=Males_data , type="percent", main = ("Infected vs Treatment"))
```

Male cases from the unfortunate covid-19 epidemic can be visual representation as just barplot's above. The blue bar's are the males who were given a plecebo or fake medications and the pink bar's are the males who were given a vaccine or an actual strong medication. At this moment we can describe that one who does not receive a vaccine or strong medication has a very higher chance of becoming ill from covid-19. That is why the light blue bars are above the pink bar, where the y axis is approximately 70 percent. Fortunately, though the next results over the x axis where it says ok are for males but who are not sick. That is why the blue and pink bars are equal, where the percent is approximately 50 on the y axis.  

## Females


## LGBTQ



## Druggies


# Overall Results and Conclusions